\textbf{NBA Party}

### 2017-04-26 init_model.rnw
p=.101 on mlb.

\paragraph{An out of the box machine learning model}
For fun, we try a simple support vector machine. We posit that whether
teams party also depends on how much time they have between games; even
when restricting our data to the twenty-four hour window after teams visit a party city, there is quite a bit of heterogeneity in when yesterday's game finished and when 
today's starts that a non-linear classifier may be able to take advantage of.

<<echo=2:5>>=
require(e1071)
svm.data <- lines[last.game.loc != team, 
                  list(hour = hour(date), last.game.time, party, 
                       loss.season, season, outcome)] %>%
    na.omit

### tune.svm(..., gamma = 2^(-5:5))
m <- svm(x = svm.data[season < 2016, -c('outcome', 'season'), with = F],
              y = svm.data[season < 2016, outcome == 'W'],
              scale = T, type = 'C-classification', probability = T, 
              kernel = 'radial')
preds <- predict(m, newdata = svm.data[season == 2016, -c('outcome', 'season'), with = F])
table(preds, svm.data[season == 2016 , outcome])
@

@@

<<echo=-1>>=
load(file = 'tmp_data/covers_lines_mlb.RData')
lines[, party := ifelse(last.game.loc != team & weekend == 1, nmusicians, 0)]

m <- glm(I(team.score > opponent.score) ~ party + I(team == location) + 
             odds + nhours.lgame + weekend + travel.dist,
         data = lines, family = 'binomial')
summary(m)
@

We note that although we don't quite achieve conventional statistical significance,
we're not that far off the mark.

@@

<<echo=F>>= 
source('functions.r')

m <- glm(I(team.score > opponent.score) ~ party,
         data = lines[year(date) < 2016], family = 'binomial')
p <- predict(m, newdata = lines[year(date) == 2016], type = 'response')

@@

sum(b, na.rm = T)

If we were to bet one hundred dollars on each game in 2016,
we would earn a profit of just less than \$400 over the course of 45 games.

>>>>

@@

<<echo=-1>>=
load(file = 'tmp_data/covers_lines_mlb.RData')
lines[, party := ifelse(team != location & weekend == 1, nmusicians, 0)]

m <- glm(I(team.score > opponent.score) ~ party + odds + I(team == location) + 
             ndays.lgame + travel.dist + weekend,
         data = lines, family = 'binomial')
summary(m)
@

We note that although we don't quite achieve conventional statistical significance,
we're not that far off the mark. The difference in significance between 
sports could possibly be explained by the heterogeneity in athleticism 
required to play basketball versus baseball, wherein the former requires far more 
non-stop sprinting. We also remark that the p-value on \texttt{party}
similar in magnitude to the p-value on our \texttt{home-team advantage} indicator,
which is an important feature as deemed by the literature.\cite{hometeamadv}

@@

<<echo=2:4>>= 
source('functions.r')
m <- glm(I(team.score > opponent.score) ~ party + odds,
         data = lines[year(date) < 2016], family = 'binomial')
p <- predict(m, newdata = lines[year(date) == 2016], type = 'response')

@@

sum(b, na.rm = T)
@

If we were to bet one hundred dollars on each game in 2016,
we would earn a profit of about \Sexpr{format(round(sum(b, na.rm=T), -1), big.mark = ',')} dollars
over the course of \Sexpr{sum(!is.na(b))} games.

\begin{thebibliography}{9}
  \bibitem{hometeamadv}
    Jeremy P. Jamieson.
    \textit{The Home Field Advantage in Athletics: A Meta-Analysis}
    \href{http://onlinelibrary.wiley.com/doi/10.1111/j.1559-1816.2010.00641.x/full}{paper}
     \end{thebibliography}
